---
title: "Hummingbird Harness Methods"
author: "Jessie Williamson"
date: "Created 10/20/2020; last revised 03/31/21"
output: html_document
---

This is a super brief script meant to make Figure S1 for Williamson and Witt (In Review), "Alightweight body harness for tracking hummingbirds" showing mass changes in Giant Hummingbirds tracked and recaptured using our harness design. 


```{R, echo=FALSE}
# I set some GLOBAL R chunk options here.
#   (to hide this message add "echo=FALSE" to the code chunk options)

knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, width = 100)
knitr::opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 6)

#knitr::opts_chunk$set(cache = TRUE, autodep=TRUE)
knitr::opts_chunk$set(cache = TRUE, autodep=TRUE)
```


# Load packages
```{R}
library(reshape)
library(reshape2)
library(plyr)
library(dplyr)
library(car)
library(GGally)
library(Hmisc)
library(gridExtra)
library(stats)
library(gplots)
library(ggplot2)
library(stats4) # Forces knitr to work when it's being wonky
library(PMCMR) #Allows Kruskal-Wallis post-hocs
library(viridis)
library(lme4)
library(sjstats)
library(MuMIn)
```


---

# Clear workspace and set WD
```{R}
rm(list=ls(all=TRUE)) # clear workspace 
setwd("/Users/Jessie/Dropbox (MSBbirds)/Rdirectory/Hummingbird_Harness_Methods")
```


# load in data 
```{R}
# Read in Table 2 with brief recapture data for Giant Hummingbirds (part of Jessie Williamson's ongoing dissertation research)
recaps <- read.csv("0_raw_data/Table2_harness_methods_recapture_data.csv", stringsAsFactors = TRUE) # n=7 geolocator individuals 
recaps.reshape <- read.csv("0_raw_data/Table2_harness_methods_recapture_data_reshape.csv", stringsAsFactors = TRUE) # for plotting, below

sl <- read.csv("Hummingbird_Survival_Estimates_forR.csv", stringsAsFactors = TRUE) # Hummingbird survival rate lit review

# Check data structure
str(recaps)
str(recaps.reshape)
recaps.reshape$nk <- as.factor(recaps.reshape$nk)
```

This is a micro-dataset, but here's what columns mean: 
NK: NK number (MSB)
Other_NK: Second NK assigned at collecting event (not always applicable; some birds have the same NK)
device_ID: Geolocator ID 
neck_loop: Harness neck loop measurement in mm 
deployment_mass: Mass taken at time of deployment (g)
recapture_mass: Mass taken at recapture (g) 
mass_diff: Difference in mass (recapture minus deployment; aka, final minus initial)
gain_loss: Direction of mass between deployment or recapture; categorical variable with 2 levels: "gain" or "loss"
capture_time: Time of capture (24 hour format)
capture_month: month of capture
notes: notes pertaining to estimations of capture time (for some birds only)

Note: I quickly converted the "recaps" data frame into long format in Excel super quickly before beginning, but this can be done in R as well (and it's better to do it in R).


# Wrangle data 
```{r}
# Rename one cumbersome column 
names(recaps)[names(recaps) == "harness_neck_loop_measurement"] <- "neck_loop" # Make this column header simpler 

# Make sure mass difference is calculated properly 
recaps$mass_diff <- recaps$recapture_mass - recaps$deployment_mass

# Reshape for making "interaction" plot below: 
# library(tidyr)
# recaps2 <- gather(recaps, "nk", "other_nk", 2:8)
# Cheated and did this quickly in Excle as a time-saver 
```

**Mass summary:** 
Loss range: -1.2 to -3.9 g loss 
Gain range: 1.0-5.8 g gain 


# Figure S1: Plot of mass change w/ NK identifier labels
```{r}
(p <- ggplot(recaps.reshape, aes(x=mass_type, y=mass, colour=gain_loss, linetype=gain_loss)) + 
      geom_point(data=recaps.reshape, aes(y=mass), size=2.8) +
      stat_smooth(data=recaps.reshape, aes(y=mass, group=nk), size=1.0, method="lm") +
      coord_cartesian(xlim = c(1.5,1.7)) + # Increase first # to reduce space between y axis and Initial mass data 
    # I originally had this set to xlim=1.45, 2.4, but right-hand side had way too much weird x-axis space
    # Setting these as is reduces blank space but increases the length of line segments connecting points, for some reason
      scale_x_discrete(labels=c("Initial \n Mass (g)", "Recapure \n Mass (g)")) + 
      annotate("text", 2.15, 19.05, label="NK252148", color = "black", size=3.5) + # NK252148 label
      annotate("text", 2.15, 21.45, label="NK252134", color = "black", size=3.5) + # NK252134 label
      annotate("text", 2.15, 22.05, label="NK252158", color = "black", size=3.5) + # NK252158 label
      annotate("text", 2.15, 23.25, label="NK252118", color = "black", size=3.5) + # NK252118 label
      annotate("text", 2.15, 23.95, label="NK252144", color = "black", size=3.5) + # NK252144 label
      annotate("text", 2.15, 24.35, label="NK252152", color = "black", size=3.5) + # NK252152 label
      annotate("text", 2.15, 24.65, label="NK252156", color = "black", size=3.5) + # NK252156 label
      # scale_shape_manual(values=c("initial"=16, "recapture"=17)) + # Manually sets shape based on geography 
      scale_color_manual(values=c(gain="darkgoldenrod2", loss="darkturquoise")) + 
      labs(x = "", y = "Mass (g)") + # Remove big x-axis label; only want "Initial mass" and "recapture mass" 
      theme_classic() +
      theme(legend.position = "none") + # Note: orange = spring, blue = winter 
      theme(axis.text.y=element_text(size=12), 
            axis.text.x=element_text(size=12, color="black"), # X-axis label font size 
            axis.title=element_text(size=14))
)
ggsave(p, filename="Figure3_HarnessMethods_Initial&RecapMassChangePlot.pdf", height=5, width=7, units="in")
```


# Compare significant differences among deployment and recapture masses (t-test and Wilcoxon signed ranks)
We'll use a paired t-test to determine whether there is a statistical difference between the means of two dependent samples (vs unpaired, which asks if there's a difference between the means of two independent samples). Paired t-test assumes relationships between groups and it does NOT assume equal variance. 
```{r}
# Take a peek at means of groups; based on these, I don't anticipate significant differences 
mean(recaps$deployment_mass)
mean(recaps$recapture_mass)

# Verify normal distributions (important because F test below is sensitive to non-normality)
qqPlot(recaps$deployment_mass) # qqPlots look fantastic 
qqPlot(recaps$recapture_mass)
hist(recaps$deployment_mass) # A little hard to see dist patterns with small sample size...qqPlot more reliable in this case. 
hist(recaps$recapture_mass)

# Quickly test for equal variances: 
var.test(recaps$deployment_mass, recaps$recapture_mass, alternative="two.sided")
# Result: p-value for the F-test is >0.05, so there is no significant difference between the two variances. 

# Test for normality 
# Note: I much prefer to evaluate normality w/ histograms and qqplots, but that's a bit difficult with n=7
shapiro.test(recaps$deployment_mass) # W = 0.84704, p-value = 0.1154
shapiro.test(recaps$recapture_mass) # W = 0.84907, p-value = 0.1205
# Result: p-value >0.05, so data are normally distributed. 

# Two sample t-test, paired 
t.summary <- t.test(recaps$deployment_mass, recaps$recapture_mass, 
                    paired=TRUE, # specifies dependence between two groups (i.e. measurements of initial and final) 
                    var.equal=TRUE, # verified w/ F-test above 
                    conf.level=0.95)
t.summary
# t=-0.0782, df=6, p=0.9402; there are no significant differences in masses between deployment and recapture 
# 95% CI: -3.229 to 3.02

# Can alternatively just run t-test on the difference of the two
# t.test(recaps$mass_diff)

# Wilcoxon signed-ranks test
# Reviewer 2 brought up the point that t-test might be affected by our small sample size, so let's compare with a 
# nonparametric test
w.summary <- wilcox.test(recaps$deployment_mass, recaps$recapture_mass, 
                    paired=TRUE, # specifies dependence between two groups (i.e. measurements of initial and final) 
                    conf.int=TRUE,
                    conf.level=0.95)
w.summary
# V = 14, p-value = 1; there are no significant differences in masses between deployment and recapture.
```


# LMER 1: Is there an effect of time of day on body mass? 
A linear mixed effects model will allow us to assess the effect of time of day on mass with a random effect of device_ID (i.e., individual), which accounts for repeated measures of individuals between deployment period (initial mass) and recapture period (recapture mass).
```{r}
# Removed 1 rows containing missing values (geom_point) --> Because no recapture mass from BJ091
# Note that recapture times for BC377, BJ090, and BJ112 were esimtated from field notes (exact capture times not recorded)

m1 <- lmer(mass ~ capture_time + (1|device_ID), data=recaps.reshape) 
summary(m1)
Anova(m1, type=3) 
# round(AIC(m1), 3) # Only useful in model comparison, which we aren't doing

# There is no effect of capture time on Giant Hummingbird mass. 


# DIAGNOSTICS
# Check assumptions w/ diagnostic plots (lm_diag_plots function doesn't work w/ S4 class)
# See good diagnostics interpretation description on pg. 292 of Discovering Statistics in R

# Residuals vs. fitted values 
plot(m1) # looks good, to be expected with extremely small sample size 

# Check distribution with normal curve overlaid
hist((resid(m1) - mean(resid(m1))) / sd(resid(m1)), freq = FALSE); curve(dnorm, add = TRUE) # looks fine

# Resid plots: No more than ~5% of cases should have abs values above 2; no more than ~1% should have abs values above 2.5
# Any absolute value above 3 is an outlier 
plot(resid(m1)) # not *ideal*, but also to be expected from small sample size... (we don't want to eliminate data!)


# QQPLOT 
qqnorm(resid(m1)) # Looks great; perhaps slight curvature at tails, but not very noticeable
qqline(resid(m1))  

# COOK'S D
plot(cooks.distance(m1)) # One influential value >1; to be conservative, we'll try rerunning the model w/out this (below)
# Any value above Cook's D of 1.0 is a highly influential value

# LEVERAGE 
plot(hatvalues(m1)) 


# Variance explained by random effect of +(1|device_ID): 
m1.indiv.var <- (2.134 / (2.134 + 5.777)); m1.indiv.var
# 26.9% of variance in model explained by differences among individals 
# This is: random effect variance / (random effect variance + residual variance)

# Calculate R^2
r2(m1) # sjstats method
r.squaredGLMM(m1) # MuMIn pseudo R-squred method, both consistent
# Marginal= variance explained by fixed effects
# Conditional = variance explained by entire model including fixed and random effects

# R^2 for full model is 0.27; R^2 for fixed effects (only time) = 0.002. 
# This makes sense given that variance explained by individual is 0.269


#####3

# RERUN MODEL WITHOUT INFLUENTIAL OBSERVATION
# Out of curiosity, let's try dropping the 1 cook's D value >1.0 (influential) to see if this affects model results
cooks.distance(m1) # This is observation 2, easily visible on our plot (BJ090 w/ low mass very late in the day)
sub <- recaps.reshape[-which(recaps.reshape$capture_time == 2115),] # very quick and dirty; just drop obs via time

m2 <- lmer(mass ~ capture_time + (1|device_ID), data=sub) # rerun model w/out influential observation
summary(m2) # Very small estimate
Anova(m2, type=3) # Capture time still does not have a significant effect on mass 
plot(cooks.distance(m2)) # Influential observation gone now 

# Since removing this influential observation doesn't substantively change our results, and
# Since we want to prioritize retaining all the observations we can, we'll report the results of the full model (m1)
# in our manuscript w/ a supplemental plot (Figure S3) of the relationship between time of day and mass. 
```


# Figure S3: Plot relationship between time of day and mass
Evidence suggests that birds may experience daily mass fluctuations (i.e., Polo & Bautista 2006). Are Giant Hummingbird masses affected by capture times? 
```{r}
(p1 <- ggplot(recaps.reshape, aes(x=capture_time, y=mass, colour=device_ID, shape=mass_type)) + 
  geom_point(size=4.5, alpha=0.8) +
  #scale_colour_viridis(discrete=TRUE) + # just use default R colors for visualization
  #geom_smooth(method = "lm", formula = y ~ x, colour="black", size=1.0) + # line used to be dark turquoise 
    # No relationship, so we don't want fit lines
  theme_classic() + 
  coord_cartesian(xlim = c(650,2200)) + # Forces y and x axes to take shape/limits you give them
  scale_x_discrete(limits=c(0700,1000,1300,1600,1900,2200), labels=c(0700,1000,1300,1600,1900, 2200)) + 
 # scale_y_discrete(limits=c(600,1,2,3,3.77), labels=c(0,10,100,1000,6000)) + # y-axis displayed in height in meters
  labs(x="Capture Time (24 hour)", 
       y="Mass (g)") +
       # theme(axis.text.x = element_text(hjust = 1)) +
  theme(legend.position = "right") + 
  annotate("text", x=2050, y=25.7, label = "italic(R) ^ 2 == 0.27", parse = TRUE) + # Add R^2 value from lmer model
  #ggtitle("A") + # Assign panel number/header; this will be (a) because first in series of 3
  theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
  theme(plot.margin = unit(c(0.2,0.1,0.0,0.2), "cm")) +  # top, right, bottom, left
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=12))
 )
ggsave(p1, filename="FigureS3_Scatterplot_EffectOfCaptureTimeOnMass.pdf", height=5, width=7, units="in")

```


# LMER 2: Is there an effect of month on body mass?
```{r}
# Quick box plot of mean mass by month; remeber only n=1 for November
p2 <- ggplot(recaps.reshape, aes(x = capture_month, y = mass))
p2 <- p2 + geom_boxplot()
p2 <- p2 + geom_point()
p2 <- p2 + stat_summary(fun.y = mean, geom = "point", shape = 18, size = 4, alpha = 0.3)
print(p2)
# This is for general assessment only; we won't include this in the supplement because not useful for many reasons
# i.e., doesn't account for individual variation, raw masses vs. masses split by capture/recapture, etc. 

# monthdata <- recaps.reshape[-which(recaps.reshape$capture_month == "November"),] # Drop Nov because n=1

m3 <- lmer(mass ~ capture_month + (1|device_ID), data=recaps.reshape) # rerun model w/out influential observation
summary(m3) # Very small estimate
Anova(m3, type=3) # Capture time still does not have a significant effect on mass 
plot(cooks.distance(m3)) 
r.squaredGLMM(m3) # Conditional R^2: 0.275; marginal R^2 = 0.223

# Variance explained by random effect of +(1|device_ID): 
m3.indiv.var <- (0.4197 / (0.4197 + 5.8751)); m3.indiv.var
# 6.7 of variance in model explained by differences among individals 
# This is: random effect variance / (random effect variance + residual variance)

library(lsmeans)
# lsmeans for capture_month; but don't report in text because not sig.  
lsmeans(m3, list(pairwise ~ capture_month           ), adjust = "tukey")
# This confirms what we know from the box plot: Dec masses are lower than January and February
# Feb masses are higher than Jan masses 
# Comparisons aren't significant. Don't interpret Nov because n=1. 

```


# Basic plot of hummingbird survival rates from the temperate zone and tropics 
```{r}
library(ggplot2)
ggplot(sl, aes(x = species, y = phi, ymin = phi-phi_se, ymax = phi+phi_se)) + 
  coord_flip() + # This flips x and y axes
    geom_linerange() + 
    geom_pointrange()

# You need to explicitly specify the width for dodging
# p + geom_linerange(aes(ymin = phi-1, ymax = y+1),
#   position = position_dodge(width = 0.9))

# Aesthetics haven't been tweaked at all! 

```

--------


# Print environment for reproducibility
```{r}
sessionInfo()
```


# End 




